Proof-Reconstruction in Type Theory for Propositional Logic

### 2017-10-21 paper/sections/proof-reconstruction.tex
[ paper ] working on corrections.

The proof-reconstruction approach proposed consists of a series
of steps similar to the workflow presented by \citeauthor{sultana2015} in
\cite{sultana2015}. This process is a translation from a source
system to a target system. In our case, the system of origin, the
automatic theorem prover, is \Metis; the target system is a proof
assistant, \Agda. We choose \Agda, but another proof assistant with the
same support of type theory and inductive types can be used.
%  (see a description of \Agda for its main features in
% Section~\ref{sec:implementation}).
The overview of the proof-reconstruction is presented in
Fig.~\ref{fig:proof-reconstruction-workflow}. The process begin with
a \TPTP file that encodes a problem in \CPL. We use this file as the
input of the \Metis prover and if the problem is a theorem, \Metis can
reply a derivation in \TSTP format of the proof.

With the \TSTP derivation from Step~2, we process the derivation with the
\Athena translator tool~\cite{Athena}.
\Athena parses the \TSTP format, analyzes the
derivation and generates a representation of the natural deduction
proof using a tree data structure (see the properties
of this tree in Section~\ref{ssec:metis-proofs}).
We perform an analysis of the proof-tree to remove some
unnecessary steps that introduce redundancies or unused input for
\Athena.

As a result of using \Athena, we get an \Agda file, the proof-term. To
generate this file, \Athena uses names of functions and theorems from
\Agda libraries:
\texttt{agda-prop}~\cite{AgdaProp}, and \texttt{agda-metis}~\cite{AgdaMetis}.

@@

% section section_name (end)
\subsection{Structural Recursion}
\label{ssec:structural-recursion}

@@

This paper adopts the notion of \emph{structurally smaller} from
\citeauthor{Abel2002}~\cite{Abel2002} to say, a recursive function is
structurally recursive if it calls itself with only
\emph{structurally smaller}
arguments.

General recursive calls can be translated into structurally recursive
calls by using for instance, the \emph{Bounded Recursion} technique
presented in~\cite{Bertot2004}. For a yet another method, we refer
the reader to \cite{Coquand1992,Abel2002,Bove2005}.
The former technique defines a new function based on the original
recursive function by adding an argument to the function.
The new argument is the \emph{bound}, a natural number given by the
function complexity. In other words, the added argument will store
the number of times the function needs to call itself to get the
expected outcome.

Henceforth, a natural number has \Nat type, and it is defined as
usual, i.e., \zero and \suc are the data constructors and some
arithmetic operations are defined: ($+,\ -,\ *$). We include syntax
sugar for $\zero$, $\suc$, $\suc(\zero)$, $\cdots$, with the
classical representation: $0$, $1$, $2$, $\cdots$ as well.

@@

\subsubsection*{A Practical Example.}
Let us consider the following example to exhibit the technique
mentioned above for an inner function present in the emulated version
of the \strip inference.

\begin{align}
  \label{eq:uh-definition}
  \begin{split}
    &\fuh_{0} : \Prop → \Prop\\
    &\fuh_{0}(φ) =\\
    &\hspace{3mm}\begin{cases}
    \fuh_{0}((φ₁ ∧ φ₂) ⇒ φ₃),
      & \text{ if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
    \fuh_{0}(φ₁ ⇒ φ₂) ∧ \fuh_{0}(φ₁ ⇒ φ₃),
      & \text{ if }φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
    φ, &\text{ otherwise.}
    \end{cases}
  \end{split}
\end{align}

From the definition above, the size of the argument, $((φ₁ ∧ φ₂) ⇒ φ₃)$,
is not structurally smaller than the input formula, $(φ₁ ⇒ (φ₂ ⇒ φ₃))$,
since $((φ₁ ∧ φ₂) ⇒ φ₃)$ is not a subformula of $(φ₁ ⇒ (φ₂ ⇒ φ₃))$.

Thus, we define the function
$uh_{1}$ by structural recursion in its second argument in
Eq.~(\ref{eq:uh-structured}).

@@

As we mentioned above, the natural number in the second argument of
the function used in Eq.~(\ref{eq:uh-structured}) stands for a
complexity measure of the function, in this case $uh_{1}$ function.

@@

\begin{align*}
  \begin{split}
  % \label{eq:rank-definition}
    &\frank : \Prop → \Nat\\
    &\frank(φ) =
    \begin{cases}
    0, &\text{for atomic }φ \\
    \frank(φ₁) + c_{¬},  &\text{ if } φ ≡ \neg φ₁ \\
    \fmax{(\frank(φ₁),\ \frank(φ₂))} + c_{\square},
      &\text{ if } φ ≡ φ₁\,\square\, φ₂.
    \end{cases}
  \end{split}
\end{align*}

We define the $\fuh_{cm}: \Prop → \NN$ function as
the complexity measure for the $\fuh_0$ function.

\begin{align*}
  \label{eq:uh-rank-definition}
  \begin{split}
      &\fuh_{cm} : \Prop → \Nat\\
      &\fuh_{cm}(φ) =\\
      &\hspace{3mm}
        \begin{cases}
        \fuh_{cm}(φ₃) + 2, &\text{ if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
        \fmax(\fuh_{cm}(φ₂),\ \fuh_{cm}(φ₃)) + 1,
          &\text{ if }  φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
        0, &\text{ otherwise.}
        \end{cases}
  \end{split}
\end{align*}

Finally, following the technique mentioned, we get a curated version of
$\fuh_{0}$ function presented in Def.~\ref{def:uh}.

\begin{definition}[uh]
  \label{def:uh}
    \begin{align*}
      \begin{split}
      &\fuh : \Prop → \Prop\\
      &\fuh(φ) = \fuh_{1}(φ, \fuh_{cm}(φ)).
      \end{split}
    \end{align*}
\end{definition}

For a more complete treatment of general recursion, and how to
define well-founded and well-structural recursive functions, we refer
the reader to Ref.~\cite{Bove2005}, where Bove and Capettra described
not only general recursion, but they also treat mutual recursive
function, among others.

@@

\subsection{Emulating \Metis' Inference Rules}
\label{ssec:emulating-inferences}

In this section, we emulate each \Metis inference rule by proving a
function with the same name, and a associated theorem.
The following pattern is used to present the formalization.
The function name written in bold print
is a \Metis inference rule in \TSTP derivations.
The same function name written in italics refers to its emulated
version.
Each emulated rule states a respective theorem that proves its soundness.
This theorem is named by using the prefix \emph{thm}
and the name of the rule.

\begin{itemize}
\item \texttt{rule} is a $n$-ary \Metis inference rule
\item $\rm{rule}$ is the emulated version function  for
the \texttt{rule} inference.
\begin{align*}
  \begin{split}
    &\rm{rule} : \Prop → \Prop → \cdots → \Prop\\
    &\rm{rule}(φ₁,\ \cdots, φₙ, ψ) =
      \begin{cases}
      ψ, &\text{ if }\rm{rule}\text{ built }ψ\text{ by applying valid}\\
         &\text{ transformations or other rules to }φ\\
      φ, &\text{ otherwise.}
      \end{cases}
  \end{split}
\end{align*}

\begin{itemize}
  \item $φᵢ$ are \emph{source} formulas from previous deducted derivations.
  \item $ψ$ is the \emph{target} formula.
  The excepted result of \texttt{rule}.
\end{itemize}
\end{itemize}

@@

\subsubsection{Stripping a Goal.}
\label{sssec:strip-a-goal}

@@

\begin{verbatim}
fof(goal, conjecture, p & r & q).
fof(subgoal_0, plain, p,
    inference(strip, [], [goal])).
fof(subgoal_1, plain, p => r,
    inference(strip, [], [goal])).
fof(subgoal_2, plain, (p & r) => q,
    inference(strip, [], [goal])).
\end{verbatim}

@@

\begin{lemma}[lem-inv-uh]
  \label{lem:uh}
  $Γ ⊢ \fuh(φ)$, then $Γ ⊢ φ$.
\end{lemma}

\begin{sketchproof} Use induction on the cases defined by the outcome of the
$\fuh$ function in Eq.~(\ref{eq:uh-structured}).
\begin{itemize}
	\item If $n = 0$, by definition we conclude $Γ ⊢ φ$.
	\item If $n = 1$, we apply the following theorem proved in~\cite{AgdaProp},
		\begin{equation*}
		\texttt{∧⇒\rm{-to-}⇒⇒}\ :\  Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃ → Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).
		\end{equation*}
  \item For $n > 1$, we use induction on the structure of the second
			  argument.
\vskip 2mm
\begin{itemize}
\item $(φ₁ ⇒ (φ₂ ⇒ φ₃))$
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \fuh(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ (φ₂ ⇒ φ₃), n-1)$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃}
  \RightLabel{∧⇒\rm{-to-}⇒⇒.}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃)}
  \end{bprooftree}
\end{equation*}

\item $(φ₁ ⇒ (φ₂ ∧ φ₃))$
\begin{equation*}
  \scalebox{0.95}{
  ($\mathcal{D}_1$)
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ \fuh(φ₁ ⇒ φ₂, n) ∧ \fuh(φ₁ ⇒ φ₃, n)$}
    \RightLabel{∧-proj₁}
    \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ φ₂, n)$}
    \RightLabel{axiom}
    \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ φ₂, n-1)$}
    \RightLabel{lem-inv-uh,}
    \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
  \end{bprooftree}
  }
\end{equation*}

\begin{equation*}
  (\mathcal{D}_2)
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ \fuh(φ₁ ⇒ φ₂) ∧ \fuh(φ₁ ⇒ φ₃)$}
    \RightLabel{∧-proj₂}
    \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ φ₃, n)$}
    \RightLabel{axiom}
    \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ φ₃, n-1)$}
    \RightLabel{lem-inv-uh.}
    \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₃$}
    \end{bprooftree}
\end{equation*}

@@

\begin{definition}[strip]
\label{def:strip}
  \begin{equation*}
    \begin{split}
    &\fstrip : \Prop → \Prop\\
    &\fstrip(φ) =
      \begin{cases}
      \fuh(\fstrip(φ₁)) ∧ \fuh(φ₁ ⇒ \fstrip(φ₂)),
      &\text{ if }φ ≡ φ₁ ∧ φ₂ \\
      \fuh(¬ φ₁ ⇒ \fstrip(φ₂)),
        & \text{ if }φ ≡ φ₁ ∨ φ₂\\
      \fuh(φ₁ ⇒ \fstrip(φ₂)),
        & \text{ if }φ ≡ φ₁ ⇒ φ₂\\
      \fuh(φ₁ ⇒ \fstrip(φ₂)) ∧ \fuh(φ₂ ⇒ \fstrip(φ₁)),
        & \text{ if }φ ≡ φ₁ ⇔ φ₂\\
      \fuh(φ₁ ⇒ \fstrip(¬ φ₂)),
        & \text{ if }φ ≡ ¬ (φ₁ ∧ φ₂)\\
      \fuh(\fstrip(¬ φ₁)) ∧\ \fuh(¬ φ₁ ⇒ \fstrip(¬ φ₂)),
        & \text{ if }φ ≡ ¬ (φ₁ ∨ φ₂)\\
      \fuh(\fstrip(φ₁)) ∧\ \fuh(φ₁ ⇒ \fstrip(¬ φ₂)),
        & \text{ if }φ ≡ ¬ (φ₁ ⇒ φ₂)\\
      \fuh(φ₁ ⇒ \fstrip(¬ φ₂)) ∧\ \fuh(¬ φ₂ ⇒ \fstrip(φ₁)),
        & \text{ if }φ ≡ ¬ (φ₁ ⇔ φ₂)\\
      \fuh(\fstrip(φ₁)),
        & \text{ if }φ ≡ ¬ ¬ φ₁\\
      ⊤,
        & \text{ if }φ ≡ ¬ ⊥\\
      ⊥,
        & \text{ if }φ ≡ ¬ ⊤.\\
      φ, &\text{ otherwise}.
      \end{cases}
    \end{split}
  \end{equation*}
\end{definition}

The definition above has termination problems since it is not
a structural recursion. Therefore, we have reformulated the equation
above to solve such an issue following the technique
described in Section~\ref{ssec:structural-recursion}.
We refer to $\fstrip$ as the function after applying the process mentioned
above here and below unless otherwise stated.

\begin{lemma}[lem-inv-strip] $Γ ⊢ \fstrip(φ)$ then $Γ ⊢ φ$.
\label{lem:lem-inv-strip}
\end{lemma}

\begin{sketchproof}
The proof is by induction on the structure of the
formula $φ$ by following the cases in Def.~\ref{def:strip}.
We present some cases but for a complete proof, we refer the reader to~\cite{AgdaMetis}.

\begin{itemize}
\item $(φ ≡ ¬ ¬ φ₁)$
\begin{equation*}
  \begin{bprooftree}
 \AxiomC{$Γ ⊢ \fstrip(φ,n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ \fuh(\fstrip(¬¬φ₁,n-1))$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ \fstrip(¬¬φ₁,n-1)$}
  \RightLabel{lem-inv-strip}
  \UnaryInfC{$Γ ⊢ ¬¬φ₁$}
  \RightLabel{axiom.}
  \UnaryInfC{$Γ ⊢ φ$}
  \end{bprooftree}
\end{equation*}

\item $(φ ≡ φ₁ ∧ φ₂)$. We get a proof for each conjunct and using the
introduction rule for conjunction connective, the result follows.

\begin{equation*}
(\mathcal{D})\hspace{3mm}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \fstrip(φ,n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁) ∧ \fuh(φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{∧-proj₁}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁)$}
  \RightLabel{lem-inv-uh.}
  \UnaryInfC{$Γ ⊢ φ₁$}
  \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{Γ ⊢ φ₁}
  \AxiomC{$Γ ⊢ \fstrip(φ,n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁) ∧ \fuh(φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{$∧$-proj$₂$}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \fstrip(φ₂,n-1)$}
  \RightLabel{$⇒$-elim}
  \BinaryInfC{$Γ ⊢ \fstrip(φ₂,n-1)$}
  \RightLabel{lem-inv-strip.}
  \UnaryInfC{$Γ ⊢ φ₂$}
  \end{bprooftree}
\end{equation*}

\item $(φ ≡ φ₁ ⇒ φ₂)$
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{}
  \RightLabel{assume φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁$}
  \AxiomC{$Γ ⊢ \fstrip(φ,n-1)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \fstrip(φ₂,n-1)$}
  \RightLabel{weaken φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁ ⇒ \fstrip(φ₂,n-1)$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ , φ₁ ⊢ \fstrip(φ₂,n-1)$}
  \RightLabel{lem-inv-strip}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₂$}
  \RightLabel{⇒-intro.}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
  \end{bprooftree}
\end{equation*}

\item $(φ ≡ ¬ (φ₁ ∨ φ₂))$. We show that $Γ ⊢ ¬ φ₁$ and $Γ ⊢ ¬ φ₂$.
The result follows by introducing the conjunction and applying
De Morgan Laws.

>>>>

The proof-reconstruction approach proposed consists of a series
of steps similar to the workflow presented by \citeauthor{sultana2015} in
\cite{sultana2015}. This process is a translation from a source
system to a target system. In our case, the system of origin, the
automatic theorem prover, is \Metis; the target system is a proof
assistant, \Agda. We choose \Agda, but another proof assistant with the
same support of type theory and inductive types could be used.
%  (see a description of \Agda for its main features in
% Section~\ref{sec:implementation}).
The overview of the proof-reconstruction is presented in
Fig.~\ref{fig:proof-reconstruction-workflow}. The process begin with
a \TPTP file that encodes a problem in \CPL. We use this file as the
input of the \Metis prover and if the problem is a theorem, \Metis can
reply a derivation of the proof in \TSTP format.

With the \TSTP derivation from Step~2, we process the derivation with the
\Athena translator tool.
\Athena parses the \TSTP format, analyzes the
derivation and generates a representation of the natural deduction
proof using a tree data structure (see the properties
of this tree in Section~\ref{ssec:metis-proofs}).
In the \Athena analysis, some unnecessary steps that introduce redundancies
and some unused input is removed from the proof-tree.

As a result of using \Athena, we get an \Agda file of the proof.
To generate this file, \Athena uses names of functions and theorems from
our \Agda libraries: \texttt{agda-prop}, and \texttt{agda-metis}.

@@

% section section_name (end)
\subsection{Structural recursion}
\label{ssec:structural-recursion}

@@

This paper adopts the notion of \emph{structurally smaller} from
\citeauthor{Abel2002}~\cite{Abel2002} to say, a recursive function is
structurally recursive if it calls itself with only
structurally smaller arguments.

General recursive calls can be translated into structurally recursive
calls by using for instance, the \emph{Bounded Recursion} technique
presented in~\cite{Bertot2004}. For a yet another method, we refer
the reader to \cite{Coquand1992,Abel2002,Bove2005}.
The former technique defines a new function based on the original
recursive function by adding an argument to the function.
The new argument is the \emph{bound}, a natural number given by the
function complexity. In other words, the added argument will represent
the number of times the function needs to call itself to get the
expected outcome~\cite{Bertot2004}.

The type of natural numbers is called \Nat type, and it is defined as
usual, i.e., \zero and \suc are the data constructors. We use names
and symbols of the arithmetic operations as usual.
We include syntax sugar for $\zero$, $\suc$, $\suc(\zero)$,
$\cdots$, with the decimal representation: $0$, $1$, $2$, $\cdots$
as well.

@@

\subsubsection*{An  example.}
Let us consider the following example to exhibit the bounded technique
mentioned above for the \fuh function presented in the version
proposed of the \Metis \strip inference.

\begin{align}
  \label{eq:uh-definition}
    &\fuh_{0} &&: \Prop → \Prop\\
    &\fuh_{0}\ (φ₁ ⇒ (φ₂ ⇒ φ₃)) &&= \fuh_{0}((φ₁ ∧ φ₂) ⇒ φ₃)\\
    &\fuh_{0}\ (φ₁ ⇒ (φ₂ ∧ φ₃)) &&= \fuh_{0}(φ₁ ⇒ φ₂) ∧ \fuh_{0}(φ₁ ⇒ φ₃)\\
    &\fuh_{0}\ φ &&= φ
\end{align}

From the definition above, the size of the argument, $(φ₁ ∧ φ₂) ⇒ φ₃$,
is not structurally smaller than the input formula, $φ₁ ⇒ (φ₂ ⇒ φ₃)$,
since $(φ₁ ∧ φ₂) ⇒ φ₃$ is not a subformula of $φ₁ ⇒ (φ₂ ⇒ φ₃)$.

Therefore, we define the function
$uh_{1}$ by structural recursion in its second argument in
Eq.~(\ref{eq:uh-structured}).

@@

As we mentioned above, the natural number in the second argument of
the function stands for a complexity measure of it.

@@

\begin{align*}
  \begin{split}
  % \label{eq:rank-definition}
    &\frank : \Prop → \Nat\\
    &\frank(φ) =
    \begin{cases}
    \frank(φ₁) + c_{¬},  &\text{ if } φ ≡ \neg φ₁ \\
    \fmax{(\frank(φ₁),\ \frank(φ₂))} + c_{\square},
      &\text{ if } φ ≡ φ₁\,\square\, φ₂.\\
    0, &\text{ otherwise}
    \end{cases}
  \end{split}
\end{align*}

We define the $\fuh_{cm}: \Prop → \Nat$ function as
the complexity measure for the $\fuh_0$ function.

\begin{align*}
  \begin{split}
      &\fuh_{cm} : \Prop → \Nat\\
      &\fuh_{cm}(φ) =\\
      &\hspace{3mm}
        \begin{cases}
        \fuh_{cm}(φ₃) + 2, &\text{ if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
        \fmax(\fuh_{cm}(φ₂),\ \fuh_{cm}(φ₃)) + 1,
          &\text{ if }  φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
        0, &\text{ otherwise.}
        \end{cases}
  \end{split}
\end{align*}

Finally, following the technique mentioned, we get a final version of
$\fuh_{0}$ function presented in Eq.~(\ref{eq:uh}).

\begin{definition}[uh]
  \label{def:uh}
    \begin{align}
      \label{eq:uh}
      \begin{split}
      &\fuh : \Prop → \Prop\\
      &\fuh(φ) = \fuh_{1}(φ, \fuh_{cm}(φ)).
      \end{split}
    \end{align}
\end{definition}

For a more complete treatment of general recursion, and how to
define well-founded and well-structural recursive functions, we refer
the reader to \citeauthor{Bove2005}~\cite{Bove2005}.

@@

\subsection{Emulating \Metis' inference rules}
\label{ssec:emulating-inferences}

In this section, we emulate each \Metis inference rule with
function and its respective theorem.
The following pattern is used to present the formalization.
The function name written in bold face
is a \Metis inference rule in \TSTP derivations.
The same function name written in roman letters refers to the proposed
version of the rule.
Each rule proposed has a respective theorem also written in roman letters
that proves its soundness.
This theorem is named by using the prefix \emph{thm}
and the name of the rule.

For example, we know the \verb!conjunct! inference rule is a
\Metis inference rule since it appears in \TSTP derivations like
the following excerpt.

\begin{verbatim}
fof(normalize_1_2, plain, (p | r),
    inference(conjunct, [], [normalize_1_1])).
\end{verbatim}

For this inference rule, we proposed the \rm{conjunct} function in
Def.~\ref{def:conjunct} and we proved the theorem \rm{thm-conjunct}
in Thm.~\ref{thm:thm-conjunct}. The pattern is the following for
a binary rule and we extend it for $n$-ary rules in a natural way.

\begin{itemize}
\item \texttt{rule} is a binary \Metis inference rule.
\item $\rm{rule}$ is the proposed function for the \texttt{rule} inference.
\begin{align*}
  \begin{split}
    &\rm{rule} : \Prop → \Prop → \Prop\\
    &\rm{rule}\ φ₁\ φ₂\ ψ\ =
      \begin{cases}
      ψ, &\text{if }\rm{rule}\text{ built }ψ\text{ by applying valid transformations}\\
         &\text{or other rules to }φ₁\text{ and }φ₂\\
      φ, &\text{otherwise.}
      \end{cases}
  \end{split}
\end{align*}

\begin{itemize}
  \item $φᵢ$ are \emph{source} formulas from previous deducted derivations.
  \item $ψ$ is the \emph{target} formula. The expected result of \texttt{rule}.
\end{itemize}
\item \rm{thm-rule} is the theorem that proves the soundness of the
$\rm{rule}$ function.
\begin{equation*}\text{If }Γ ⊢ φ\text{ then }Γ ⊢ rule\ φ.\end{equation*}

\end{itemize}

@@

\subsubsection{Stripping a goal.}
\label{sssec:strip-a-goal}

@@

\begin{verbatim}
fof(goal, conjecture, ((p & r) & q)).
fof(subgoal_0, plain, p,
    inference(strip, [], [goal])).
fof(subgoal_1, plain, p => r,
    inference(strip, [], [goal])).
fof(subgoal_2, plain, (p & r) => q,
    inference(strip, [], [goal])).
\end{verbatim}

@@

\begin{lemma}[lem-inv-uh]
  \label{lem:uh}
  If $Γ ⊢ \fuh\ φ$ then $Γ ⊢ φ$.
\end{lemma}

\begin{sketchproof}
Use induction on the cases defined by the outcome of the
$\fuh$ function in Eq.~(\ref{eq:uh-structured}).
\begin{itemize}
	\item If $n = 0$, by definition we conclude $Γ ⊢ φ$.
	\item If $n = 1$, we apply the following theorem proved in~\cite{AgdaProp},
		\begin{equation*}
		\texttt{∧⇒\rm{-to-}⇒⇒}\ :\  Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃ → Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).
		\end{equation*}
  \item For $n > 1$, we use induction on the structure of the second
			  argument.
\vskip 2mm
\begin{itemize}

\item Case $φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)$.
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \fuh(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ (φ₂ ⇒ φ₃), n-1)$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃}
  \RightLabel{∧⇒\rm{-to-}⇒⇒.}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃)}
  \end{bprooftree}
\end{equation*}

\item Case $φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)$.
\begin{equation*}
  \scalebox{0.95}{
  ($\mathcal{D}_1$)
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ \fuh(φ₁ ⇒ φ₂, n) ∧ \fuh(φ₁ ⇒ φ₃, n)$}
    \RightLabel{∧-proj₁}
    \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ φ₂, n)$}
    \RightLabel{by def.}
    \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ φ₂, n-1)$}
    \RightLabel{lem-inv-uh,}
    \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
  \end{bprooftree}
  }
\end{equation*}

\begin{equation*}
  (\mathcal{D}_2)
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ \fuh(φ₁ ⇒ φ₂) ∧ \fuh(φ₁ ⇒ φ₃)$}
    \RightLabel{∧-proj₂}
    \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ φ₃, n)$}
    \RightLabel{by def.}
    \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ φ₃, n-1)$}
    \RightLabel{lem-inv-uh.}
    \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₃$}
    \end{bprooftree}
\end{equation*}

@@

\begin{definition}[strip]
\label{def:strip}
  \begin{align}
  \label{eq:strip}
    &\fstrip &&: \Prop → \Nat → \Prop\\
    &\fstrip\ (φ₁ ∧ φ₂)\ n     &&= \fuh\ (\fstrip\ φ₁) ∧ \fuh\ (φ₁ ⇒ \fstrip\ φ₂)\\
    &\fstrip\ (φ₁ ∨ φ₂)\ n     &&= \fuh\ (¬ φ₁ ⇒ \fstrip\ φ₂)\\
    &\fstrip\ (φ₁ ⇒ φ₂)\ n     &&= \fuh\ (φ₁ ⇒ \fstrip\ φ₂)\\
    &\fstrip\ (φ₁ ⇔ φ₂)\ n     &&= \fuh\ φ₁ ⇒ \fstrip\ φ₂) ∧ \fuh\ φ₂ ⇒ \fstrip φ₁)\\
    &\fstrip\ (¬ (φ₁ ∧ φ₂))\ n &&= \fuh\ φ₁ ⇒ \fstrip(¬ φ₂))\\
    &\fstrip\ (¬ (φ₁ ∨ φ₂))\ n &&= \fuh\ \fstrip(¬ φ₁)) ∧\ \fuh ¬ φ₁ ⇒ \fstrip(¬ φ₂))\\
    &\fstrip\ (¬ (φ₁ ⇒ φ₂))\ n &&= \fuh\ \fstrip(φ₁)) ∧\ \fuh φ₁ ⇒ \fstrip(¬ φ₂))\\
    &\fstrip\ (¬ (φ₁ ⇔ φ₂))\ n &&= \fuh\ φ₁ ⇒ \fstrip(¬ φ₂)) ∧\ \fuh ¬ φ₂ ⇒ \fstrip(φ₁))\\
    &\fstrip\ (¬ ¬ φ₁)\ n      &&= \fuh\ \fstrip(φ₁))\\
    &\fstrip\ (¬ ⊥)\ n         &&= ⊤\\
    &\fstrip\ (¬ ⊤.)\ n        &&= ⊥\\
    &\fstrip\ φ\ n             &&= φ
  \end{align}
\end{definition}

% The definition above has termination problems since it is not
% a structural recursion. Therefore, we have reformulated the equation
% above to solve such an issue following the technique
% described in Section~\ref{ssec:structural-recursion}.
% We refer to $\fstrip$ as the function after applying the process mentioned
% above here and below unless otherwise stated.

\begin{lemma}[lem-inv-strip]
If $Γ ⊢ \fstrip\ φ$ then $Γ ⊢ φ$.
\label{lem:lem-inv-strip}
\end{lemma}

\begin{sketchproof}
The proof is by induction on the structure of the
formula~$φ$ by following the cases in Eq.~(\ref{eq:strip}).
We present some cases but for a complete proof, we refer the reader to~\cite{AgdaMetis}.

\begin{itemize}
\item Case $φ ≡ ¬ ¬ φ₁$.
\begin{equation*}
  \begin{bprooftree}
 \AxiomC{$Γ ⊢ \fstrip(¬¬φ₁,n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh(\fstrip(φ₁,n-1))$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ \fstrip(φ₁,n-1)$}
  \RightLabel{lem-inv-strip}
  \UnaryInfC{$Γ ⊢ φ₁$}
  \end{bprooftree}
\end{equation*}

\item Case $φ ≡ φ₁ ∧ φ₂$. We get a proof for each conjunct and using the
introduction rule for conjunction connective, the result follows.

\begin{equation*}
(\mathcal{D})\hspace{3mm}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \fstrip(φ,n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁) ∧ \fuh(φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{∧-proj₁}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁)$}
  \RightLabel{lem-inv-uh.}
  \UnaryInfC{$Γ ⊢ φ₁$}
  \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{Γ ⊢ φ₁}
  \AxiomC{$Γ ⊢ \fstrip(φ,n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁) ∧ \fuh(φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{$∧$-proj$₂$}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \fstrip(φ₂,n-1)$}
  \RightLabel{$⇒$-elim}
  \BinaryInfC{$Γ ⊢ \fstrip(φ₂,n-1)$}
  \RightLabel{lem-inv-strip.}
  \UnaryInfC{$Γ ⊢ φ₂$}
  \end{bprooftree}
\end{equation*}

\item Case $φ ≡ φ₁ ⇒ φ₂$.
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{}
  \RightLabel{assume φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁$}
  \AxiomC{$Γ ⊢ \fstrip(φ,n-1)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \fstrip(φ₂,n-1)$}
  \RightLabel{weaken φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁ ⇒ \fstrip(φ₂,n-1)$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ , φ₁ ⊢ \fstrip(φ₂,n-1)$}
  \RightLabel{lem-inv-strip}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₂$}
  \RightLabel{⇒-intro.}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
  \end{bprooftree}
\end{equation*}

\item Case $φ ≡ ¬ (φ₁ ∨ φ₂)$. We show that $Γ ⊢ ¬ φ₁$ and $Γ ⊢ ¬ φ₂$.
The result follows by introducing the conjunction and applying
De Morgan Laws.
